#Analysis 3
#Getting only the curated Genes

The purpose of this is to use all genes in creating the SOM, using only leaf development genes.  These come from Yasu's curated Gene list. 


##Required Libraries

```{r}
library(ggplot2)
library(reshape)
library(kohonen)
library(plyr)
```

Upload Yasu's curated gene list

```{r}
countData <- read.csv("../data/normalized_read_count.csv")
geneList1 <- read.csv("../../../06diffGeneExp/analysis/indvGenes/yasuCuratedGenes/pnas.1402835111.sd06.csv")
geneList2 <- read.csv("../../../06diffGeneExp/analysis/indvGenes/yasuCuratedGenes/pnas.1402835111.sd07.csv")

#Merge the Curated Gene Lists Together
genes2 <- geneList1[,1]
genes1 <- geneList2[,1]

genesOfInterest1 <- levels(genes1)
genesOfInterest2 <- levels(genes2)

genesOfInterest <- c(genesOfInterest1, genesOfInterest2)
head(genesOfInterest)
```


```{r}
#Melt Data
countData <- melt(countData)
colnames(countData) <- c("gene", "sample", "count")


#set genotype

countData$genotype <- ifelse(grepl("wt", countData$sample, ignore.case = T), "wt", 
         ifelse(grepl("tf2", countData$sample, ignore.case = T), "tf2", "unknown"))

#set tissue

countData$tissue <- ifelse(grepl("other", countData$sample, ignore.case = T), "other", 
         ifelse(grepl("mbr", countData$sample, ignore.case = T), "mbr", "unknown"))

#Set Region
countData$region <- ifelse(grepl("a", countData$sample, ignore.case = T), "A", 
         ifelse(grepl("c", countData$sample, ignore.case = T), "C", "B"))

#Set type

countData$type <- paste(countData$region, countData$tissue,  sep = "")
```

Intitialize dataframe to fit in all the visualization information. 

```{r}
largeGeneList <- data.frame(t(rep(NA,7)))
colnames(largeGeneList) <- c("type", "genotype", "N", "mean", "sd", "se", "gene")
head(largeGeneList)
```
##Making the data table for Visualization 

This loop 
(Takes about 20 minutes) 
```{r, tidy = FALSE}
for(GENE in genesOfInterest) {
 
  if(length(grep(GENE, countData$gene)) < 1){ #this is just making sure that the curated
    next; 
    }
      
  geneData <- subset(countData, grepl(GENE, countData$gene))
   
  sumGraph <- ddply(geneData, c("type", "genotype"), summarise,
               N    = length(count),
               mean = mean(count),
               sd   = sd(count),
               se   = sd / sqrt(N))
  
  sumGraph$gene <- GENE
  
largeGeneList  <- rbind(largeGeneList, sumGraph) #bind together all the new rows per loop. 
}

#make for use
finalList <- largeGeneList
finalList <- finalList[-1,] #remove the first row

write.table(finalList, file = "../data/analysis03_GeneList.csv", sep = ",")

#get the log2 of mean. 

finalList$log2Mean <- log2(finalList$mean) 
plot(finalList$mean) #Why are there negative numbers here?
finalList <- subset(finalList, log2Mean > 0 ) #fix that problem, but must go back to figure out why
```


Self Organizing Maps: Ciera LCM 
========================================================

The goal of this analysis is to find genes that have co-expression patterns of 

1. curated leaf gene list. Mean, Rep,
2. most differentiated genes between tissue. 

##Required Libraries

```{r}
library(ggplot2)
library(reshape)
library(kohonen)
```

##1.pca.R

The first step is to get it into the right format. First column being the genes, while the subsequent columns are the different libraries (type).

```{r}
leafGenes <- read.csv("../data/analysis02_GeneList.csv")
head(leafGenes)
leafGenes <- leafGenes[c(7, 1, 4)] #keep only needed columns (gene, type, mean)
summary(leafGenes)
head(leafGenes)

#Change from long to wide data format
leafGenes.long <- cast(leafGenes, gene ~ type, value.var = mean, fun.aggregate = "mean")  #why did I have to specify "mean" here? Are there duplicates of types? Double check later. 

leafGenes.long <- as.data.frame(leafGenes.long) #or does not keep column names during transformation. 
names(leafGenes.long)

scale_data <- as.matrix(t(scale(t(leafGenes.long[c(2:7)])))) 

head(scale_data)

#Principle Component Analysis
pca <- prcomp(scale_data, scale=TRUE) 

summary(pca) 
pca.scores <- data.frame(pca$x)

data.val <- cbind(leafGenes.long, scale_data, pca.scores) 
head(data.val)
```


##Visualizing the PCA

There is not clear clustering in these genes. I don't think using all the genes is an appropriate way to find clear patterns in this data.  There is too much noise. I will proceed anyway, to explore how this effects SOM results.  

```{r}
p <- ggplot(data.val, aes(PC1, PC2))
p + geom_point(alpha = 0.3)
```

##1. Self Organizing Map

First I am going to include all the different tissue regions. I have set the somgrid to 3,3, which shows seperation. 

```{r}
names(data.val)

som.data <- as.matrix(data.val[,c(2:7)]) #This is where you set how many libraries are involved. 
head(som.data)
set.seed(2) # This does not appear to influence the outcome very much 


som <- som(data=som.data, somgrid(3,3,"hexagonal"))
summary(som)
```


The training progress seems appropriate.

```{r}
plot(som, type ="changes")
```

The codes plot is showing some groups of genes are similarly expressed in certain tissues. I will have to look into GO expression to understand what these genes are exactly. The leaf specific genes are patterned in a way in which the C region (base) does not have many of the leaf specific genes. 

```{r}
plot(som, type = "codes")
```

The counts plot is worrisome to me. This is showing how many genes are clustered in each category.  It shows that most are evenly distibuted, which, to me, shows that the seperation could be random, there is not clear delimiting factors involved in seperating these clusters. 

```{r}
plot(som, type = "counts")
```

As with counts, the "quality" is showing how much these clusters seperate, which in this case shows even seperation.  I would feel more comfortable with the results if they showed a variation in seperation because that would suggest gene expresssion values showed enough variation in pattern to influence variation in cluster seperation. Although in the Kohonen_vignette.pdf, they explain "The right plot (quality) shows the mean distance of objects, mapped to a particular unit, to the codebook vector of that unit. **A good mapping should show small distances everywhere in the map**." All of the distances are small so maybe it is good?  Need to look into more. 
```{r}
plot(som, type = "quality")
```

Self-Organizing Map - Supervised Mapping 
=========================



