#Getting the most sig. differentiated genes between tissue

##Purpose
This should be the last analaysis of the original SOM.  This analysis is a follow-up to `lcmSOM_analysis1c_080514.Rmd`, where the original scaling will be done specific to genotype.  

##Required Libraries

```{r}
library(ggplot2)
library(reshape)
library(kohonen)
```

Self Organizing Maps 
========================================================

##1.pca.R

First read in file that came from mostSigDEgenes.Rmd.  This is a list of genes from all DE analysis in WT. They were all cancatenated, then duplicate genes were removed.  In addition the mean was calculated from the replicates of each type. 

The first step is to get it into the right format. First column being the genes, while the subsequent columns are the different libraries (type).

```{r}
mostDEgenes <- read.csv("../data/allGeneListBothGenotypes_analysis5b.csv")
head(mostDEgenes)

mostDEgenes <- mostDEgenes[c(7, 1, 2, 4)] #keep only needed columns (gene, type, mean)

#Change from long to wide data format
mostDEgene.long <- cast(mostDEgenes, genotype + gene ~ type, value.var = mean, fun.aggregate = "mean")  #why did I have to specify "mean" here? Are there duplicates of types? Double check later. 
mostDEgene.long <- as.data.frame(mostDEgene.long)  
head(mostDEgene.long)
```

At this point I am going to subset on genotype and scale seperatley before adding them back together. **THIS  IS THE ONLY PART DIFFERENT FROM analysis`lcmSOM_analysis1c_080514.Rmd`**

```{r}
wt <- subset(mostDEgene.long, genotype == "wt")
tf2 <- subset(mostDEgene.long, genotype == "tf2")

scale_data.wt <- as.matrix(t(scale(t(wt[c(3:8)]))))#transformation.
scale_data.tf2 <- as.matrix(t(scale(t(tf2[c(3:8)]))))#transformation.
?scale
scale_data <- rbind(scale_data.wt, scale_data.tf2)
```

```{r}
#Principle Component Analysis
pca <- prcomp(scale_data, scale=TRUE) 

summary(pca) 

pca.scores <- data.frame(pca$x)

data.val <- cbind(mostDEgene.long, scale_data, pca.scores) 
head(data.val)
```

##Visualizing the PCA

Looks to be three major clusters. 

```{r}
p <- ggplot(data.val, aes(PC1, PC2)) 
p + geom_point(alpha = .6)
```

------------

*I am skipping the large Map and going straight to the small*

2. Self Organizing Map- Small (3,2)
-----------------------------

The size of the map is something that may cause differences in the genes that are clustered.  Using a small map size (3,2), I found they cluster in according to tissue type. See below.

```{r}
names(data.val)
som.data <- as.matrix(data.val[,c(9:14)])
head(som.data)
set.seed(5)

som <- som(data=som.data, somgrid(3,2,"hexagonal")) # This is where you change the size of the map

summary(som)
```

###Training Plot ("changes") - Small

This shows a hundred iterations. Training plateaus at around 20 iterations steeply.  Is this a problem?

```{r}
plot(som, type ="changes")
```

###Code Plot - Small

Here with the small map, each tissue has a tissue specific cluster. 

```{r}
plot(som, type = "codes")
```

###Count Plot - Small

This tells you how many genes are in each of the clusters. 

```{r}
plot(som, type = "counts")
```

###Distance Neighbour Plot- Small

This is sometimes called the "U-Matrix", it can help identify further clustering. Areas of low neighbour distance indicate groups of nodes that are similar and the further apart nodes indicate natural "borders" in the map. 

```{r}
plot(som, type="dist.neighbours")
```

###Heatmaps - Small 

This shows the distribution of each type of tissue.  This doesn't really work too well when the the map is so small. Bother is the only tissue type that contributes to two clusters. 

```{r}
head(som$codes)
som$data <- data.frame(som$data) #changed to dataframe to extract column names easier. 

#This is just a loop that plots the distrinution of each tissue type across the map. 
for (i in 1:6){
  plot(som, type = "property", property = som$codes[,i], main=names(som$data)[i])
  print(plot)
  }
```

Output
```{r}
data.val.small <- cbind(data.val,som$unit.classif,som$distances)
head(data.val.small)
write.table(data.val.small, file="../data/analysis1.som.data.small.ALLD.txt")
#Make sure that there is just one of each value som$unit.classif and distances column. 
names(data.val.small)

summary(data.val.small)
```

#Visualize back to PC Space

```{r}
plot.data <- read.table("../data/analysis1.som.data.small.ALLD.txt",header=TRUE)
names(plot.data)

p <- ggplot(plot.data, aes(PC1, PC2, colour=factor(som.unit.classif))) #use unit.classif for smaller dataset
p + geom_point(alpha = .6) + facet_grid(.~genotype) + theme_bw()
```

##References

1. [R self Organizing map tutorial](http://www.r-bloggers.com/self-organising-maps-for-customer-segmentation-using-r/)
