#Super SOM

##Purpose

The purpose of this analysis is to make a superSOM.  

```{r}
library(ggplot2)
library(reshape)
library(plyr)
library(kohonen)
```

Read in files and format data from raw files
=============================================

Create a list of all DE analysis:

###Count Data
```{r, eval=FALSE}
#Read in count data
countData <- read.csv("../data/normalized_read_count.csv")
#Melt count data
countData <- melt(countData)
colnames(countData) <- c("gene", "sample", "count")

#set genotype

countData$genotype <- ifelse(grepl("wt", countData$sample, ignore.case = T), "wt", 
         ifelse(grepl("tf2", countData$sample, ignore.case = T), "tf2", "unknown"))

#set tissue

countData$tissue <- ifelse(grepl("other", countData$sample, ignore.case = T), "other", 
         ifelse(grepl("mbr", countData$sample, ignore.case = T), "mbr", "unknown"))

#Set Region
countData$region <- ifelse(grepl("a", countData$sample, ignore.case = T), "A", 
         ifelse(grepl("c", countData$sample, ignore.case = T), "C", "B"))

#Set type

countData$type <- paste(countData$region, countData$tissue,  sep = "")
head(countData)
```

###Most Significantly DE genes
```{r, eval = FALSE}

#you need to reset working directory to allSigGenes for easier import of list of significant genes
setwd("~/Dropbox/Research/LCMProject/08SOM/lcmSOM/data/allSigGenes/")

#makes a list of all files in directory
file_list <- list.files()

#Creates one list from all the files in directory
for (file in file_list){
       
  # if the merged dataset doesn't exist, create it
  if (!exists("dataset")){
    dataset <- read.table(file, header=TRUE, sep="\t")
  }
   
  # if the merged dataset does exist, append to it
  if (exists("dataset")){
    temp_dataset <-read.table(file, header=TRUE, sep="\t")
    dataset<-rbind(dataset, temp_dataset)
    rm(temp_dataset)
  }
 
}

names(dataset)
dim(dataset)

allGenes<- dataset

#recieve just the list
allGenesITAG <- allGenes[,1]
length(allGenesITAG)

#Remove duplicates
allGenesITAG <- unique(allGenesITAG)

length(allGenesITAG)

#make an empty table to hold all the genes
allGeneList <- data.frame(t(rep(NA,7)))
colnames(allGeneList) <- c("type", "genotype", "N", "mean", "sd", "se", "gene")
allGeneList <- allGeneList[-1,] #remove first row
head(allGeneList)
```

#Loop together all relevent gene information.

```{r, eval = FALSE}
for(GENE in allGenesITAG) {
 
  if(length(grep(GENE, countData$gene)) < 1){ #this is just making sure that the list of sig genes
    next; 
    }
      
  geneData <- subset(countData, grepl(GENE, countData$gene))
   
  sumGraph <- ddply(geneData, c("type", "genotype"), summarise,
               N    = length(count),
               mean = mean(count),
               sd   = sd(count),
               se   = sd / sqrt(N))
  
  sumGraph$gene <- GENE
  
allGeneList  <- rbind(allGeneList, sumGraph) #bind together all the new rows per loop. 
}
dim(allGeneList)
head(allGeneList)

#There is only wt here!
allGeneList$genotype <- as.factor(allGeneList$genotype)
levels(allGeneList$genotype)

#at this point set working directory back to source

write.table(allGeneList, file = "../data/allGeneListBothGenotypes_analysis5b.csv", sep = ",")
```

##1.pca.R


```{r}

mostDEgenes <- read.csv("../data/allGeneListBothGenotypes_analysis5b.csv")

mostDEgenes <- mostDEgenes[c(7, 2, 1, 4)] #keep only needed columns (gene, genotype, type, mean)

head(mostDEgenes)

#Change from long to wide data format
mostDEgene.long <- cast(mostDEgenes, genotype + gene ~ type, value.var = mean, fun.aggregate = "mean")  #why did I have to specify "mean" here? Are there duplicates of types? Double check later. 

head(mostDEgene.long)

mostDEgene.long <- as.data.frame(mostDEgene.long) #transformation. 
names(mostDEgene.long)


scale_data <- as.matrix(t(scale(t(mostDEgene.long[c(3:8)]))))
head(scale_data) 
#Principle Component Analysis
pca <- prcomp(scale_data, scale=TRUE) 

summary(pca) 

pca.scores <- data.frame(pca$x)

data.val <- cbind(mostDEgene.long, scale_data, pca.scores) 

head(data.val)
```


##Visualizing the PCA

Looks to be three major clusters. 

```{r}
p <- ggplot(data.val, aes(PC1, PC2)) 
p + geom_point()
```


##SuperSOM

```{r}
set.seed(6)
names(data.val)

superSomData <- data.val[,c(1:2,9:14)]

#it seems like I should have had the genotype information in the header in the first place so now I am going to take the long way around to fix that.  The better idea would have to have done this *before* the seprating them.  Do this later. 

tf2 <- subset(superSomData, genotype == "tf2", select = 3:8)
wt <- subset(superSomData, genotype == "wt", select = 3:8)

colnames(tf2) <- c("tf2Ambr", "tf2Aother", "tf2Bmbr", "tf2Bother", "tf2Cmbr", "tf2Cother") 
colnames(wt) <- c("wtAmbr", "wtAother", "wtBmbr", "wtBother", "wtCmbr", "wtCother")

wt <- as.matrix(wt)
tf2 <- as.matrix(tf2)

#I may need to scale seperatly? Does it make a difference?
#sc.m82 <- t(scale(t(m.m82)))
#sc.penn <- t(scale(t(m.penn)))

all.data <- list(wt,tf2)

ssom <- supersom(all.data, somgrid(3, 2, "hexagonal"),weights=c(0.5,0.5)) 

summary(ssom)
par(mfrow = c(3, 2))
plot(ssom, type ="changes")
plot(ssom, type = "codes")
plot(ssom, type = "counts")
plot(ssom, type = "quality")

data.val <- cbind(data.val,tf2,wt,ssom$unit.classif,ssom$distances)

head(data.val)

write.table(data.val, file="../data/ssom.data.analysis5.txt")
```

This is setting the larger clusters that incorporate multiple clusters. 

```{r}

## use hierarchical clustering to cluster the codebook vectors
som_cluster <- cutree(hclust(dist(ssom$codes)), 5)
# plot these results:
plot(som, type="mapping", bgcol = som_cluster, main = "Clusters") 
add.cluster.boundaries(som, som_cluster) 

# I want to attach the hierchal cluster to the larger dataset data.val.
som_clusterKey <- data.frame(som_cluster)
som_clusterKey$unit.classif <- c(1:36)

data.val <- cbind(data.val,som$unit.classif,som$distances)

#Merge data.val with som_clusterKey
##change data.val to match som_cluster key 
names(data.val)[20] <- "unit.classif"

data.val <- merge(data.val, som_clusterKey, by.x = "unit.classif" ) #ignore warning, this is what you want.  You are essentially filling in the empties with the value from som_clusterKey

write.table(data.val, file="../data/analysis1.som.data.txt")
#Make sure that there is just one of each value som$unit.classif and distances column. 
names(data.val)
```




###Visualization 

```{r}
plot.data <- read.table("../data/ssom.data.analysis5.txt",header=TRUE)
names(plot.data)
dim(plot.data)

p <- ggplot(plot.data, aes(PC1, PC2, colour=factor(ssom.unit.classif))) #notice I am using som_cluster and not unit.classif, because this is the larger SOM
p + geom_point() + theme_bw()
```