#Analysis 7 - Getting a closer look at the clusters from the orignal clusters. 

##Purpose

To get start to understand the differences in GO categories between the clusters. This is the data from analysis1D. 

###Part 1

This will look into the number of genes that are the same between the clusters and the genotypes. With some basic visualization.

Required Libraries

```{r}
library(VennDiagram)
library(ggplot2)
library(reshape)
library(kohonen)
library(goseq)
library(GO.db)
```

###Visualize by Cluster
Read in data used for GO enrichment analysis

```{r}
geneLength <- read.csv("../../../07GO_enrichment/requisiteData/normalized_genes_length.csv")
cate <- read.table("../../../07GO_enrichment/requisiteData/melted.GOTable.txt",header=TRUE)
```

Read in data produced from analysis1D.
```{r}
plot.data <- read.table("../data/analysis1.som.data.small.ALLD.txt",header=TRUE)
names(plot.data)
```


##Cluster Specific analysis
Now I want to take a look at what are is going on exactly in these clusters. The clusters start with the bottom left, which is cluster number 1.

This is a function that makes a boxplot showing the transformed values of expression in the clusters. 

```{r}
#clusterVis Function
#displays transformed data in a box plot and 
clusterVis <- function(clustNum){
  
  sub_cluster <- subset(plot.data, som.unit.classif==clustNum)
  sub_data <- sub_cluster[,9:14] # just the sample types
  m.data <- melt(sub_data) 
  p <- ggplot(m.data, aes(x=variable, y=value))
  p + geom_point(alpha=0.5, position="jitter", size=1) + geom_boxplot(alpha=0.75, outlier.size=0) 
}
```

Number of genes function, which gives you some basics about the clusters between 

```{r}
clusterNum <- function(clustNum){

  sub_cluster <- subset(plot.data, som.unit.classif==clustNum)
  print(paste("total number of genes in sub cluster is ", 
              nrow(sub_cluster)
              )
        )

  scwt <- subset(sub_cluster, genotype == "wt")
  print(paste("total number of genes in wt cluster is ", 
              nrow(scwt)
              )
        )

  sctf2 <- subset(sub_cluster, genotype == "tf2")
  print(paste("total number of genes in tf2 cluster is ", 
              nrow(sctf2)
              )
        )
  
  scIntersect <- as.data.frame(intersect(scwt$gene, sctf2$gene))
   print(paste("There are", 
               length(intersect(scwt$gene, sctf2$gene)), 
               " that are the same between wt and tf2"
               )
         )
  
  ##Venn Diagram part
  grid.newpage()
  venn.plot <- draw.pairwise.venn(area1 = nrow(scwt),
                              area2    = nrow(sctf2),
                              cross.area = length(intersect(scwt$gene, sctf2$gene)),
                              scaled       = F,
                              category     = c("Wildtype", "tf2"),
                              fill         = c("blue", "red"),
                                alpha        = 0.3,
                                lty          = "blank",
                                cex          = 2,
                                cat.cex      = 2,
                                cat.pos      = c(315, 25),
                                cat.dist     = 0.09,
                                cat.just     = list(c(-1, -1), c(1, 1)),
                                ext.pos      = 30,
                                ext.dist     = -0.05,
                                ext.length   = 0.85)
  grid.draw(venn.plot)

}


```


```{r}

 
clusterGO <- function(clustNum){
##GO Enrichment on the catergories

#we need to first get the data in the right format.
#First get the list of ITAG, 

#sub_cluster
sub_cluster <- subset(plot.data, som.unit.classif==clustNum)
scwt <- subset(sub_cluster, genotype == "wt")
sctf2 <- subset(sub_cluster, genotype == "tf2")
scIntersect <- as.data.frame(intersect(scwt$gene, sctf2$gene))

itag.sc <- as.data.frame(sub_cluster$gene) 
colnames(itag.sc)[1] <- "itag"
itag.sc$sc <- 1  

#scwt
itag.scwt <- as.data.frame(scwt$gene) 
colnames(itag.scwt)[1] <- "itag"
itag.scwt$wt <- 1  

#sctf2
itag.sctf2 <- as.data.frame(sctf2$gene) 
colnames(itag.sctf2)[1] <- "itag"
itag.sctf2$tf2 <- 1  

#Intersect
itag.scIntersect <- as.data.frame(scIntersect[1]) 
colnames(itag.scIntersect)[1] <- "itag"
itag.scIntersect$intersect <- 1 

#Merge all by itag
ITAGmerge <- merge(itag.scIntersect, itag.scwt, by = "itag", all= TRUE)
ITAGmerge <- merge(ITAGmerge, itag.sctf2, by = "itag", all = TRUE)
matrixGO <- merge(ITAGmerge, geneLength, by = "itag", all = TRUE)
matrixGO[is.na(matrixGO)] <- 0
pat <- matrixGO

#Now that we have the data in the right format we can proceed with GO enrichment.

#First specify vector to loop over for each column

sigType <- c("intersect", "wt", "tf2")

  for(type in sigType) {
  
    genes = as.integer(pat[,type])
    names(genes) = pat$itag
    table(genes)
    length(genes)
  
    pwf = nullp(genes,bias.data=pat$length)
  
    GO.wall = goseq(pwf,gene2cat = cate)
    head(GO.wall)
  
  #This is going to correct for multiple testing.  You can specify the p-value cut-off of GO categories you are interested.
  
    enriched.GO = GO.wall$category[p.adjust(GO.wall$over_represented_pvalue, method = "BH") < 0.05]
  
    enriched.GO
  
    my.GO <- as.character(enriched.GO)
    my.GO.table <- Term(my.GO)
    my.GO.table
    t <- as.matrix(my.GO.table)
  
    print(type) #this is for the knitr document
    print(t) #this is for the knitr document
  }
}
```

vennDiagram Function:  


###Cluster 1

Sub cluster 1 is defined by up regulation of genes in Bmbr, which is the early leaflet region of the terminal leaflet. 

```{r}
clusterVis(1)
clusterNum(1)
clusterGO(1)
```

###Cluster 2

Sub cluster 2 is defined by up regulation of genes in Cmbr, which is the base "marginal blastozone" region, which should be the most pluripotent in WT. 

```{r}
clusterVis(2)
clusterNum(2)
clusterGO(2)
```

###Cluster 3

This cluster is specific to Cother, which is specific to the rachis region at the base. 

```{r}
clusterVis(3)
clusterNum(3)
clusterGO(3)
```

###Cluster 4 

This cluster has genes that are preferentially up-regulated in Ambr, which is the tip most region that becomes the terminal leaflet.  This is the terminal leaflet blade region 
```{r}
clusterVis(4)
clusterNum(4)
clusterGO(4)
```

###Cluster 5 

The cluster ihas genes that are preferentially up-regulated in Aother, which is the rachis region at the tip; what will eventually become the midvien of the terminal leaflet.

```{r}
clusterVis(5)
clusterNum(5)
clusterGO(5)
```

###Cluster  6

The cluster ihas genes that are preferentially up-regulated in Bother, which is the rachis region at site of leaflet initiation.

```{r}
clusterVis(6)
clusterNum(6)
clusterGO(6)
```
