#Analysis 4 - Top 25% of coefficient of variation

In this analysis I am checking to see if using the top 25% of genes based on co-efficient of variation will give me similar results as Analysis 1, which was based on all genes from DE analysis between all tissues in WT. 

##1. Get the co-efficient of variation. 
```{r}
countData <- read.csv("../data/normalized_read_count.csv")
head(countData)

names(countData)

countData$tf2ambr <- rowMeans(subset(countData[2:5]))
countData$tf2aother <- rowMeans(subset(countData[6:9]))
countData$tf2bmbr <- rowMeans(subset(countData[10:12]))
countData$tf2bother <- rowMeans(subset(countData[13:16]))
countData$tf2cmbr <- rowMeans(subset(countData[17:20]))
countData$tf2cother <- rowMeans(subset(countData[21:24]))

countData$wtambr <- rowMeans(subset(countData[25:27]))
countData$wtaother <- rowMeans(subset(countData[28:32]))
countData$wtbmbr <- rowMeans(subset(countData[33:36]))
countData$wtbother <- rowMeans(subset(countData[37:40]))
countData$wtcmbr <- rowMeans(subset(countData[41:46]))
countData$wtcother <- rowMeans(subset(countData[47:50]))
dim(countData)
ave <- subset(countData[50:61])
ave$sd <- apply(ave,1,function(d)sd(d))
ave$average <- rowMeans(subset(ave[1:12]))
ave$cv <- ave$sd / ave$average
dim(ave)

countData <- cbind(countData, ave[,13:15])
names(countData)
```

```{r}
quantile(countData$cv)
countData[is.na(countData)] <- 0
subCountData <- subset(countData, cv > 0.61357492)
names(subCountData)
head(subCountData)
allGenes25 <- subCountData[,c(1,50:64)]

colnames(allGenes25)[1]<-"gene"
```

##2.pca

```{r}
write.csv(allGenes25, "../data/analysis4.top25.csv")
scale_data <- as.matrix(t(scale(t(allGenes25[c(8:13)])))) 

#Principle Component Analysis
pca <- prcomp(scale_data, scale=TRUE) 

summary(pca) 
pca.scores <- data.frame(pca$x)

data.val.allGenes25 <- cbind(allGenes25, scale_data, pca.scores) 
head(data.val.allGenes25)
```

##Visualizing the PCA
The PCA looks very similar. 

```{r}
p <- ggplot(data.val.allGenes25, aes(PC1, PC2))
p + geom_point()
```

##What are the differences in genes?

What is the difference between genes between allGeneList, which is from analysis one and only includes genes from the list of wt. 

```{r}
mostDEgenesWT <- read.csv("../data/allGeneList.csv") #just WT
head(mostDEgenesWT)

mostDEgenesWT <- mostDEgenesWT[c(7, 1, 4)] #keep only needed columns (gene, type, mean)

#Change from long to wide data format
mostDEgenesWT.long <- cast(mostDEgenesWT, gene ~ type, value.var = mean, fun.aggregate = "mean") 


#must cast from short to long version for comparisons. 
head(mostDEgenesWT.long)
head(allGenes)
dim(mostDEgenesWT.long) #2249
dim(allGenes25) #4618 #25% cv

```

There are more genes in the top 25%.
```{r}
length(intersect(mostDEgenesWT.long$gene, allGenes25$gene)) #1500
length(setdiff(mostDEgenesWT.long$gene, allGenes25$gene)) #749
length(setdiff(allGenes25$gene,mostDEgenesWT.long$gene)) #3118
```

There are 2,249 genes in WT only DE genes and 4,618 in the top 25% co-efficient of variation.

There are 1500 genes that are the same between 25% co-efficient of variation and the top genes in DE expression. There are 749 unique genes in DE expression list and 3118 in the top 25% co-efficient of variation.  

How would this affect the SOMs?  It would cast a wider net when performing the SOMs.  But the difference between them is somewhat worrisome.  There should not be genes found in mostDEgenesWT that are not found in the top 25% from co-efficient of variation.  

Caveats:  
This is not including the DE genes from tf2, which IS taken into account in the 25% co-efficient of variation.

##Comparing list of all genes

The best comparison would be all the genes in all list of DE, as I did in Analysis5, this list was found reading in all of the genes of all the 27 different DE expression analysis. 

```{r}
#Read in total list and make long
mostDEgenesALL <- read.csv("../data/allGeneListBothGenotypes_analysis5.csv")

mostDEgenesALL <- mostDEgenesALL[c(7, 1, 4)] #keep only needed columns (gene, type, mean)
head(mostDEgenesALL)
#Change from long to wide data format
mostDEgenesALL.long <- cast(mostDEgenesALL, gene ~ type, value.var = mean, fun.aggregate = "mean") 
head(mostDEgenesALL.long)
dim(mostDEgenesALL.long) #3,574
```

```{r}
length(intersect(allGenes25$gene, mostDEgenesALL.long$gene)) #2355
length(setdiff(allGenes25$gene, mostDEgenesALL.long$gene)) #2263
length(setdiff(mostDEgenesALL.long$gene, allGenes25$gene)) #1219
```

There are 2355 similar genes between all DE genes between all tissues and tissue types (mostDEgenesALL),  There are 2263 genes that are unique to top 25% of co-efficient of variatation, and 1219 unique to mostDEgenesALL.  Can I just use all?  Can I add the 2,263 genes found from top 25% of co-efficient of variation to my list? Ask this question in tomato group. 

I think I am going to go ahead and start using the list of unique genesbetween all DE genes between all tissues and tissue types as found in analysis5. 

##Curated Genes?

Make `data.val.allGenes25` match curated gene list. 

```{r}
data.val.allGenes25$gene <- gsub("^(.*)[.].*", "\\1", data.val.allGenes25$gene)
data.val.allGenes25$gene <- gsub("^(.*)[.].*", "\\1", data.val.allGenes25$gene)
```

```{r}
geneList1 <- read.csv("../../../06diffGeneExp/analysis/indvGenes/yasuCuratedGenes/pnas.1402835111.sd06.csv")

#isolate the genes
genesOfInterest <- geneList1[,c(1,3)]
colnames(genesOfInterest) <- c("gene", "name")
names(genesOfInterest) #check
```


```{r}
#This is most stupid way around assigning if a gene is a curated gene! 
#Figure out more elegent way.
data.val.allGenes25$curated <- match(data.val.allGenes25$gene, genesOfInterest$gene)
data.val.allGenes25$curated <- gsub("[[:digit:]]+", "yes", data.val.allGenes25$curated)
data.val.allGenes25$curated[is.na(data.val.allGenes25$curated)] <- "no"
```


###Visualize

```{r}
p <- ggplot(data.val.allGenes25, aes(PC1, PC2, color = curated)) 
p + geom_point(size=I(5), alpha = 0.6) +
  scale_colour_manual(values=c("#cccccc", "#33ff33")) + 
  theme_bw() + facet_grid(.~curated)
```

###How many curated genes are in my normalized read count?

```{r}
head(countData)

countData$X <- gsub("^(.*)[.].*", "\\1", countData$X)
countData$X <- gsub("^(.*)[.].*", "\\1", countData$X)

intersect(genesOfInterest$gene, countData$X) #There are only 217 leaf curated genes in the normalized read counts. 
intersect(genesOfInterest$gene, data.val.allGenes25$gene) # There are only 77 which are in the top 25%
#There are 51 which overlap with the DE genes.  
```

